\vskip-0.6in Endogenous Politics and the Design of Trade Institutions

### 2016-12-10 draft.tex
make minor changes from read through sec:rigid and sec:escape

When political pressure is endogenous, in place of taking expectations over a range of probabilistically-determined levels of exogenous political pressure, we can backward induct to determine the lobby's effort decision and the government's optimal choice of tariff caps in anticipation of the lobby's behavior.

@@

We know that the lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding. The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$. The government is unable to use a weak binding to encourage extra lobbying in the case with only one import-competing lobby under consideration here.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}

@@

Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage according to the chosen tariff cap $\tau^R_{W,e}$. The governments get to choose the $\tau^R_{W,e}$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.

Notice that the maximand in Expression~\ref{exp:1} is exactly the same as that in this third-stage maximization problem except it adds  foreign government welfare. This implies that $\tau^R_{W,e}$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments do not want as much lobbying as possible. To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.}

@@

On the other hand, if the level of the optimal weak binding $\tau^R_{W,e}$ is strictly positive, the weak binding is at least in part serving to generate political involvement---most often thought of as campaign contributions---that the government finds beneficial from its politically-motivated point of view. The government could have chosen a cap of zero and eliminated lobbying activity altogether. In fact, any strong binding accomplishes this very feat, but is also \textit{unable} to generate lobbying activity because of its effect on lobbying incentives. Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.}

@@

In order to establish an equilibrium, we must ensure that both the on-schedule (static) incentive constraint and the off-schedule (repeated-game) incentive constraint is satisfied. We know that the former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here. Following most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternatives that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.}
	%see KRW, Buzard 2014 for alternatives

@@

When $\ga$ is exogenous, the optimal trade agreement tariff is the same under perfect external enforcement and self-enforcement because the change in enforcement conditions cannot affect the realization of $\ga$. We will see that moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem. Moving to the repeated-game environment with self enforcement does not alter the governments costs and benefits so that the government's optimal trade agreement tariff is invariant to the enforcement conditions. This creates the possibility that the politically-optimal trade agreement tariff may not be self-enforcing.

@@

To see this, start by noticing that the lefthand side of Expression~\ref{exp:govincent} in increasing in $\de$, while the righthand side is constant in $\de$. That is, as the government becomes more patient, it is easier to satisfy the government's constraint. Equivalently, it requires a larger $e^B$ to violate the constraint.

@@

\begin{proposition}
  The presence of the lobby as a repeated-game player may imply that the politically optimal self-enforcing trade agreement tariff is strictly greater than $\tau^R_{W,e}$. 
\end{proposition}

@@

From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section~\ref{sec:objfcn}, the governments' ex-ante welfare is reduced by excess lobbying.

@@

\subsection{An Escape Clause For Endogenous Politics}
\label{sec:ECendog}
In this world that is far from first best, I next demonstrate how a WTO-style escape clause can allow flexibility in the face of fundamental shocks $s$ while discouraging lobbyists from using the shock as cover for rent-seeking behavior. The timeline is as follows:
\begin{enumerate}
	\item The government sets the trade agreement tariff $\tau^F_{W,e}$ cooperatively. I assume this is a weak binding and add the subscript $F$ to denote the environment with flexibility to deal with shocks.
	\item The exogenous shock $s$ is realized. Its level is assumed costlessly verifiable for simplicity.
	\item The lobby chooses its effort level $e^F$.
	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau(s)$ and $\tau^B(\ga(s,e^F))$.
\end{enumerate}
Here I take the higher, `escape' tariff to be continuous instead of set at a given binding. If the government applies $\tau^F_{W,e}$, implementation of the agreement goes on as before. Consistent with the WTO Agreement on Safeguards, the government can apply the tariff that is necessary to remedy the injury caused by the shock, $\tau(s)$, with no retaliation for three periods. If the government applies a tariff that is higher than either $\tau^F_{W,e}$, $\tau(s)$, it will choose its myopic best response to the pressure it experiences. I label this tariff $\tau^B(\ga(s,e^F))$, where $e^F$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.

@@

Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may chooses to set the trade agreement tariff in the presence of an escape clause $\left(\tau^F_{W,e}\right)$ so that in some cases the lobby's incentive constraint is violated. This would lead to the legislature's constraint being violated as well, creating a dispute.

@@

It still does not seem feasible to provide this flexibility at precisely the level that the shock impacts the government's preferences $\ga(s)$. What's more, this may provide an answer to why Safeguard measures have fallen out of use. If the real purpose is to provide an outlet for political pressure, I have demonstrated above that the current WTO-style Safeguards rules seem designed explicitly to discourage rent-seeking behavior. Thus if other measures such as Anti-Dumping Duties are more susceptible to manipulation, it seems natural that lobbies' efforts would migrate there.

>>>>

When political pressure is endogenous, in place of taking expectations over a range of probabilistically-determined levels of exogenous political pressure, we backward induct to determine the lobby's effort decision and the government's optimal choice of tariff caps in anticipation of the lobby's behavior.

@@

The lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding. The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$. The government is unable to use a weak binding to encourage extra lobbying when only import-competing interests are represented by a lobby.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}

@@

Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage. The governments get to choose the $\tau^R_{W,e}$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.

Notice that the maximand in Expression~\ref{exp:1} is exactly the same as that in this third-stage maximization problem except it adds  foreign government welfare. This implies that $\tau^R_{W,e}$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments want as much lobbying as possible. To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.\label{fn:tot}}

@@

On the other hand, if the level of the optimal weak binding $\tau^R_{W,e}$ is strictly positive, the weak binding is at least in part serving to generate political involvement---most often thought of as campaign contributions---that the government finds beneficial from its politically-motivated point of view. The government could have chosen a cap of zero and eliminated lobbying activity altogether. In fact, any strong binding accomplishes this very feat, but is also \textit{unable} to generate lobbying activity because of its effect on lobbying incentives. Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.\label{fn:mrc1}}

@@

In order to establish an equilibrium, we must ensure that both the on-schedule (static) incentive constraint and the off-schedule (repeated-game) incentive constraint is satisfied. We know that the former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here. Following most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternative that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.\label{fn:krw}}
	%see KRW, Buzard 2014 for alternatives

@@

When $\ga$ is exogenous, the optimal trade agreement tariff is the same under perfect external enforcement and self-enforcement because the change in enforcement conditions cannot affect the realization of $\ga$. We will see that moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem. Moving to the repeated-game environment with self enforcement does not alter the government's costs and benefits so that the government's optimal trade agreement tariff is invariant to the enforcement conditions. This creates the possibility that the politically-optimal trade agreement tariff may not be self-enforcing.

@@

To see this, start by noticing that the lefthand side of Expression~\ref{exp:govincent} is increasing in $\de$, while the righthand side is constant in $\de$. That is, as the government becomes more patient, it is easier to satisfy the government's constraint. Equivalently, it requires a larger $e^B$ to violate the constraint.

@@

\begin{proposition}
  The presence of the lobby as a repeated-game player may imply that the politically optimal self-enforcing trade agreement tariff is strictly greater than $\tau^R_{W,e}$.
	\label{res:repeated}
\end{proposition}

@@

From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section~\ref{sec:objfcn}, the governments' ex-ante welfare is reduced by excess lobbying. However, the problem is not interesting if government welfare is everywhere increasing in lobbying effort because the lower binding will be set to the prohibitive tariff in this environment with no counter-lobbying by export groups.

@@

\subsection{An Escape Clause For Endogenous Politics}
\label{sec:ECendog}
In this world that is far from first best, I next demonstrate how a WTO-style escape clause can allow flexibility in the face of fundamental shocks $s$ while discouraging lobbyists from using the shock as cover for rent-seeking behavior. The timeline is as follows:
\begin{enumerate}
	\item The government sets the trade agreement tariff $\tau^F_{W,e}$ cooperatively. I assume this is a weak binding and add the subscript $F$ to denote the environment with flexibility to deal with shocks.
	\item The exogenous shock $s$ is realized. Its level is assumed costlessly verifiable for simplicity.
	\item The lobby chooses its effort level $e^F$.
	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau(s)$ and $\tau^B(\ga(s,e^F))$.
\end{enumerate}
Here I take the higher, `escape' tariff to be continuous instead of set at a given binding. If the government applies $\tau^F_{W,e}$, implementation of the agreement goes on as before. Consistent with the WTO Agreement on Safeguards, the government can apply the tariff that is necessary to remedy the economic injury---not necessarily equal to the political injury---caused by the shock, $\tau(s)$, with no retaliation for three periods. If the government applies a tariff that is higher than either $\tau^F_{W,e}$ or $\tau(s)$, it will choose its myopic best response to the pressure it experiences. I label this tariff $\tau^B(\ga(s,e^F))$, where $e^F$ must be high enough to give the government incentive to break the trade agreement in the absence of external enforcement.

@@

Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau^F_{W,e}\right)$ so that in some cases the lobby's incentive constraint is violated. This would lead to the legislature's constraint being violated as well, creating a dispute.

@@

It still does not seem feasible to provide this flexibility at precisely the level that the shock impacts the government's preferences $\ga(s)$ if the WTO does not know the mapping from $s$ to $\ga$. What's more, this may provide an answer to why Safeguard measures have fallen out of use. If the real purpose is to provide an outlet for political pressure, I have demonstrated above that the current WTO-style Safeguards rules seem designed explicitly to discourage rent-seeking behavior. Thus if other measures such as Anti-Dumping Duties are more susceptible to manipulation, it seems natural that lobbies' efforts would migrate there.

### 2016-12-10 to-do.tex
make minor changes from read through sec:rigid and sec:escape

		\item Review appendix of JIE$\_$revision to see what connections I want to make

@@

@@

\section{Misc.}
\begin{itemize}
	\item Define $\ga$ before using it
	\item If add back in assumption 1, add Ethier footnote
	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
	\item Ben: I'm wondering how you view the time-frame over which your model is set.  On one hand a specific-factors set-up is normally associated with the short-run.  But on the other hand you have an infinite time horizon which suggests very long run.
	\item Clarify information structure
\end{itemize}

>>>>

@@

				\item Footnote fn:mrc1 is good place to start to put references to MRC 2007

@@

\section{Misc.}
\begin{itemize}
	\item Define $\ga$ before using it
	\item If add back in assumption 1, add Ethier footnote
	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
	\item Ben: I'm wondering how you view the time-frame over which your model is set.  On one hand a specific-factors set-up is normally associated with the short-run.  But on the other hand you have an infinite time horizon which suggests very long run.
	\item Clarify information structure
	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together better
	\item Add ref to JIE R$\&$R to fn:krw ?
	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
\end{itemize}
